### 2020-07-12 paper/dcpo_demsupport_appendix.Rmd

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "[dcpo_demsupport] Appendix"
thanks: "Corresponding author: [yuehong-tai@uiowa.edu](mailto:yuehong-tai@uiowa.edu). Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: ""
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
nocite: |
  @Solt2020a
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300)

# If you haven't install `DCPOtools`
# remotes::install_github("fsolt/DCPOtools")

if (!require(pacman)) install.packages("pacman")
library(pacman)

# load all the packages you will use below 
p_load(
  dataverse, # data scraping
  DCPOtools, # analysis
  plm,
  huxtable, kableExtra, modelsummary, # tabulation
  latex2exp, # visualization
  rstan, # Bayesian estimation
  tidyverse # data wrangling
)

# Functions preload
set.seed(313)
```

```{r claassen_input}
dcpo_input_raw1 <- read_csv("data/dcpo_input_raw.csv", col_types = "cdcddcd") %>% 
  filter(!(str_detect(survey, "army_wvs") & # WVS obs identified as problematic by Claassen 
                 ((country=="Albania" & year==1998) |
                      (country=="Indonesia" & (year==2001 | year==2006)) |
                      (country=="Iran" & year==2000) |
                      (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                      (country=="Vietnam" & year==2001)) |
                 (str_detect(item, "strong_wvs") &
                      ((country=="Egypt" & year==2012) |
                           (country=="Iran" & (year==2000 | year==2007)) | # 2005 in Claassen
                           (country=="India") |
                           (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                           (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
                           (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
                           (country=="Vietnam" & year==2001))) |
                 (country %in% c("Puerto Rico", "Northern Ireland", 
                                 "SrpSka Republic", "Hong Kong SAR China")))) %>%
  with_min_yrs(2)

claassen_input <- DCPOtools::format_claassen(dcpo_input_raw1)
save(claassen_input, file = here::here("data", "claassen_input.rda"))
```

```{r claassen_replication_input}
# Note that before running `dataverse::get_file()` below, one should set their personal token and server in their system environment first: 
# Sys.setenv("DATAVERSE_KEY" = "exampleToken")
# Sys.setenv("DATAVERSE_SERVER" = "dataverse.harvard.edu")
# These values can be set to persist across R sessions using `usethis::edit_r_environ()`

supdem <- dataverse::get_file("supdem raw survey marginals.tab", "doi:10.7910/DVN/HWLW0J") %>% 
    read_csv(col_types = "cdcddcdc")

supdem_cy <- supdem %>%                                             # 1390 obs
    janitor::clean_names() %>% 
    mutate(old_country = country,
           country = countrycode::countrycode(old_country, "country.name", "country.name")) %>%
    dplyr::select(country, year, project) %>% 
    unique()

claassen_input_cy <- read_csv("data/dcpo_input_raw.csv",
                              col_types = "cdcddcd") %>%           # 1813 obs
  mutate(p_dcpo = str_extract(survey, "^[a-z]+"), 
         project = case_when(p_dcpo == "afrob" ~ "afb",
                             p_dcpo == "amb" ~ "lapop",
                             p_dcpo == "arabb" ~ "arb",
                             p_dcpo == "asiab" ~ "asb",
                             p_dcpo == "asianb" ~ "asnb",
                             p_dcpo == "neb" ~ "ndb",
                             p_dcpo == "sasianb" ~ "sab",
                             TRUE ~ p_dcpo),
         y_dcpo = year) %>%
  dplyr::select(country, year, y_dcpo, survey, project) %>% 
  unique()

no_problems <- inner_join(supdem_cy, claassen_input_cy)             # 1297 obs

needed <- anti_join(supdem_cy, claassen_input_cy)                   # 93 obs

available <- anti_join(claassen_input_cy, supdem_cy)                # 516 obs

year_fixes <- left_join(needed, available, by = c("country", "project")) %>% # 89 obs
    mutate(diff = year.x - year.y) %>% 
    group_by(country, project, year.x) %>% 
    mutate(closest_to_claassen = min(abs(diff))) %>% 
    ungroup() %>% 
    group_by(country, project, year.y) %>% 
    mutate(closest_to_dcpo = min(abs(diff))) %>% 
    ungroup() %>% 
    filter(closest_to_claassen == abs(diff) & closest_to_dcpo == abs(diff) & abs(diff) <= 3) %>% 
    filter(!(country == "Egypt" & year.x == 2014 & survey == "afrob5")) # double match (it's really afrob6)

cys_crosswalk <- year_fixes %>% 
    select(country, y_dcpo, y_claassen = year.x, survey)

missing_cyps <- anti_join(needed, year_fixes,  by = c("country", "year" = "year.x", "project")) # 4 obs; listed in issue #5

cys_to_drop <- anti_join(available, year_fixes, by = c("country", "year" = "year.y", "project")) %>% # 427 obs
    select(-y_dcpo)

claassen_replication_input_raw1 <- read_csv("data/claassen_input_raw.csv", col_types = "cdcddcd") %>% 
  filter(!(str_detect(item, "army_wvs") & # WVS obs identified as problematic by Claassen 
    ((country=="Albania" & year==1998) |
         (country=="Indonesia" & (year==2001 | year==2006)) |
         (country=="Iran" & year==2000) |
         (country=="Pakistan" & (year==1997 | year==2001)) | #in DCPO,it is 1997 instead of 1996
         (country=="Vietnam" & year==2001)) |
    (str_detect(item, "strong_wvs") &
         ((country=="Egypt" & year==2012) |
              (country=="Iran" & (year==2000 | year==2007)) | #in DCPO, the year is 2007 rather than 2005
              (country=="India") |
              (country=="Pakistan" & (year==1997 | year==2001)) | #in DCPO,it is 1997 instead of 1996
              (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
              (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
              (country=="Vietnam" & year==2001))))) %>% 
  anti_join(cys_to_drop, by = c("country", "year", "survey")) %>% # surveys unused by Claassen
  left_join(cys_crosswalk, by = c("country", "year" = "y_dcpo", "survey")) %>% 
  mutate(year = if_else(!is.na(y_claassen), y_claassen, year)) %>% # use Claassen's year codings
  mutate(item = if_else(item == "strong_amb_1" & year == 2004, "strong_amb_2", item)) %>% # items conflated in amb_combo file
  with_min_yrs(2)

claassen_replication_input <- DCPOtools::format_claassen(claassen_input_raw1)
save(claassen_replication_input, file = here::here("data", "claassen_replication_input.rda"))

dcpo_replication_input <- DCPOtools::format_dcpo(claassen_input_raw1,
                                                 scale_q = "church_lb",
                                                 scale_cp = 2)
save(dcpo_replication_input, file = here::here("data", "dcpo_replication_input.rda"))
```

```{r claassen_m5, eval=FALSE}
claassen_m5 <- rstan::stan(file = 'R/argon/dcpo_demsupport/R/supdem.stan.mod5.stan',
                           data = claassen_input,
                           iter = 3000,
                           chains = 4,
                           cores = 4,
                           thin = 3,
                           pars = c("mu_lambda", "sigma_lambda", "sigma_delta", "sigma_theta", "phi", "lambda", "delta", "theta", "x_pred","log_lik"),
                           control = list(adapt_delta=0.99, stepsize=0.02, max_treedepth=11))

save(claassen_m5, file = "data/claassen_m5.rda")

claassen_m5_theta <- rstan::extract(claassen_m5, pars = "theta")

save(claassen_m5_theta, file = "data/claassen_m5_theta.rda")

claassen_m5_x_pred <- rstan::extract(claassen_m5, pars = "x_pred")

save(claassen_input, claassen_m5_x_pred, file = "data/claassen_m5_x_pred.rda")
```

```{r comparison_table}
load(here::here("data", "claassen_m5_x_pred.rda"))

ivt_claassen <- DCPOtools::internal_validation_tests(claassen_input, claassen_m5_x_pred, "claassen")

load(here::here("data", "dcpo_replication_y_r_pred.rda"))

ivt_dcpo <- DCPOtools::internal_validation_tests(dcpo_replication_input, dcpo_replication_y_r_pred)
# see https://github.com/fsolt/dcpo_article/blob/master/R/validation_table.R for more to add when we have it

```

```{r claassen_input}
claassen_input_raw1 <- read_csv("data/dcpo_input_raw.csv", col_types = "cdcddcd") %>% 
  filter(!(str_detect(survey, "army_wvs") & # WVS obs identified as problematic by Claassen 
                 ((country=="Albania" & year==1998) |
                      (country=="Indonesia" & (year==2001 | year==2006)) |
                      (country=="Iran" & year==2000) |
                      (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                      (country=="Vietnam" & year==2001)) |
                 (str_detect(item, "strong_wvs") &
                      ((country=="Egypt" & year==2012) |
                           (country=="Iran" & (year==2000 | year==2007)) | # 2005 in Claassen
                           (country=="India") |
                           (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                           (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
                           (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
                           (country=="Vietnam" & year==2001))) |
                 (country %in% c("Puerto Rico", "Northern Ireland", 
                                 "SrpSka Republic", "Hong Kong SAR China")))) %>%
  with_min_yrs(2)

claassen_input <- DCPOtools::format_claassen(claassen_input_raw1)
save(claassen_input, file = here::here("data", "claassen_input.rda"))

```

```{r claassenWUncertainty}
load("../data/claassen_replication_input.rda")
load("../data/claassen_m5_theta.rda")

ls_year <- 1988:2017
ls_country <- claassen_replication_input$data$country %>% unique

df_thetaDraw <- map_df(seq(ls_year), # year2
       function(oneYear) {
         map_df(seq(ls_country),
                function(oneCountry) {
                  df_temp <-
                    sample(claassen_m5_theta$theta[, oneYear, oneCountry],
                           size = 100, #draw 100 from the distribution
                           replace = TRUE) %>%
                    matrix(ncol = 100) %>%
                    as_tibble()
                  df_temp$year <- ls_year[oneYear]
                  df_temp$country <- ls_country[oneCountry]
                  return(df_temp)
                })
       }) %>% 
  pivot_longer(
    cols = starts_with("V"),
    names_to = "draw",
    values_to = "theta"
  ) %>% 
  group_split(draw)

df_controls <- read_csv("../data/Support_democracy_ajps.csv")

df_thetaDraw <- map(df_thetaDraw, function(oneDraw){
  left_join(oneDraw, df_controls, by = c("country" = "Country", "year" = "Year")) %>% 
    pdata.frame(index = c("country", "year"))
})

result_m5O <- plm(Libdem_VD ~ plm::lag(Libdem_VD, 1:2) + plm::lag(SupDem_trim, 1) + plm::lag(lnGDP_imp, 1) + plm::lag(GDP_imp_grth, 1) + plm::lag(Libdem_regUN, 1) + plm::lag(Pr_Muslim, 1) + plm::lag(Res_cp_WDI_di, 1), pdata.frame(df_controls, index = c("Country", "Year")), model="pooling")

result_m5N <- map(df_thetaDraw, function(aDataset){
  plm(Libdem_VD ~ plm::lag(Libdem_VD, 1:2) + plm::lag(theta, 1) + plm::lag(lnGDP_imp, 1) + plm::lag(GDP_imp_grth, 1) + plm::lag(Libdem_regUN, 1) + plm::lag(Pr_Muslim, 1) + plm::lag(Res_cp_WDI_di, 1), model="pooling", data = aDataset)
})

vcovBK_se <- function(x){
  vcovBK(x, cluster = "time") %>% 
    sqrt(diag(.))}

result_betas_m5 <- MIextract(result_m5, fun = coef)
result_vars_m5 <- MIextract(result_m5, fun = vcovBK_time)

summary(result_m5O)
summary(MIcombine(results = result_betas_m5, 
                  variance = result_vars_m5))
```

# Conclusion

\pagebreak