### 2023-06-06 replication_materials/data_replication/data_replication.Rmd
update data_replication

---
title: "Data Replication Script for Democracy, Public Support, and Measurement Uncertainty"
output:
    - pdf_document
date: "2023-06-01"
---

This file generates the latent variable estimates of support for democracy employed in the article, "Democracy, Public Support, and Measurement Uncertainty."  It then combines these estimates with data on the control variables to create the analysis datasets.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE) # comment this line to run all code
# knitr::opts_chunk$set(echo = TRUE)             # uncomment this line to run all code

library(tidyverse)
library(cmdstanr)
library(DCPOtools)
library(here)
library(furrr)
```

# Data Replication Files
There are 14 data files in the `data_replication` directory:

| Type         | Filename 
|--------------|---------------------------------
| Input        | mood_dem.csv
| Intermediate | claassen_input_raw.rds
| Intermediate | claassen_replication_input.rda
| Intermediate | exp_claassen_input.rds
| Intermediate | claassen_replication_theta.rds
| Intermediate | exp_theta.rds
| Intermediate | exp_dcpo_theta.rds
| Input        | cls_ajps_cntrl_list.rds
| Output       | correct_cls_ajps.rda
| Input        | cls_apsr_cntrl_list.rds
| Output       | correct_cls_apsr.rda
| Input        | exp_ajps_cntrl_list.rds
| Output       | expcor_cls_ajps.rda
| Output       | dcpo_ajps.rda
| Input        | exp_apsr_cntrl_list.rds
| Output       | expcor_cls_apsr.rda
| Output       | dcpo_apsr.rda

"Input" files are exogenous.  "Intermediate" files are generated by input files (and other intermediate files) and are used to generate output files; all of these processes are included in this document.  "Output" files are generated by input and intermediate files, and are used in the analyses scripted in the `dcpo_demsupport.Rmd` file (as is the `exp_claassen_input.rds` intermediate file).

# Source Data for Latent Variable Estimates

This section generates the source data needed to estimate the latent variable.
The first chunk starts from the original surveys; it generates the file `claassen_input_raw.rds`.  Collecting all of the needed survey datasets is labor-intensive (see https://github.com/fsolt/DCPOtools), so by default this code chunk is not evaluated and the next chunk simply uses the `claassen_input_raw.rds` file included in the replication materials.

```{r claassen_input_raw, eval=FALSE}
claassen_input_raw <- DCPOtools:::claassen_setup(
    vars = read_csv(here("data_replication",
                         "mood_dem.csv"),
                    col_types = "cccccc"),
    file = here::here("data_replication",
                      "claassen_input_raw.rds"),
    datapath = "../../data/dcpo_surveys")
```

The object `claassen_input_raw` contains many more observations than the dataset employed in Claassen (2020a,b), and that dataset misidentifies the survey year in ~8% the observations it does include.  The next chunk reproduces the sample of countries and years included in those data and saves it as `claassen_replication_input.rds`.  The expanded dataset labeled "more data" in our work is saved as `exp_claassen_input.rds`.

```{r claassen_replication_input}
claassen_input_raw <- rio::import(here::here("data_replication",
                      "claassen_input_raw.rds"))

dem <- readRDS(here("data_replication", "exp_ajps_cntrl_list.rds")) %>% 
    first() %>% 
    transmute(country, year, dem = as.numeric(Regime_VD > 1))

claassen_input_raw$scared <- claassen_input_raw %>% 
    pluck("lower") %>% 
    left_join(dem, by = join_by(country, year)) %>% 
    mutate(r = case_when(r == -1 & dem == 0 ~ 999, # code non-respondents in 
                         TRUE ~ r)) %>%  # autocracies as supporting democracy
    select(-dem)

remove_bad_items <- function(x) {
    x %>% 
    filter(!(str_detect(survey, "army_wvs") & # WVS obs identified as problematic by Claassen 
                     ((country=="Albania" & year==1998) |
                          (country=="Indonesia" & (year==2001 | year==2006)) |
                          (country=="Iran" & year==2000) |
                          (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                          (country=="Vietnam" & year==2001)) |
                     (str_detect(item, "strong_wvs") &
                          ((country=="Egypt" & year==2012) |
                               (country=="Iran" & (year==2000 | year==2007)) | # 2005 in Claassen
                               (country=="India") |
                               (country=="Pakistan" & (year==1997 | year==2001)) | # 1996 in Claassen
                               (country=="Kyrgyzstan" & (year==2003 | year==2011)) |
                               (country=="Romania" & (year==1998 | year==2005 | year==2012)) |
                               (country=="Vietnam" & year==2001))) |
                     (country %in% c("Puerto Rico", "Northern Ireland", 
                                     "SrpSka Republic", "Hong Kong SAR China"))))
}

claassen_input_raw1 <- map(claassen_input_raw, function(df) { 
    df %>%
        remove_bad_items() %>%
        with_min_yrs(2)
})

exp_claassen_input <- format_claassen(claassen_input_raw1)

rio::export(exp_claassen_input, file = here::here("data_replication",
                                           "exp_claassen_input.rds"))

dir.create("data", showWarnings = FALSE)
rio::export(exp_claassen_input, file = here::here("data",
                                           "exp_claassen_input.rds"))

# Note that before running `dataverse::get_file()` below, one should set their personal token and server in their system environment first: 
# Sys.setenv("DATAVERSE_KEY" = "exampleToken")
# Sys.setenv("DATAVERSE_SERVER" = "dataverse.harvard.edu")
# These values can be set to persist across R sessions using `usethis::edit_r_environ()`

if (!file.exists(here("data_replication", "supdem raw survey marginals.tab"))) {
    tempfile <- dataverse::get_file_by_doi("doi:10.7910/DVN/HWLW0J/RA8IJC") # AJPS replication file

    writeBin(tempfile, here("data_replication",
                                  "supdem raw survey marginals.tab"))
}

supdem <- read_csv(here::here("data_replication", "supdem raw survey marginals.tab"),
                   col_types = "cdcddcdc")

supdem1 <- supdem %>%                                             # 1390 obs
    janitor::clean_names() %>% 
    mutate(old_country = country,
           country = countrycode::countrycode(old_country, "country.name", "country.name"),
           i_claassen0 = tolower(item),
           i_claassen = case_when(str_detect(i_claassen0, "strong_arb") ~ "strong_arb",
                                  str_detect(i_claassen0, "strong_lapop") ~ "strong_lapop",
                                  str_detect(i_claassen0, "strong_pew") ~ "strong_pew",
                                  TRUE ~ i_claassen0),
           year = if_else(c_abb == "INS" & year == 2010 & i_claassen0 == "strong_pew",
                          2011,
                          year)) %>% 
    with_min_yrs(2)

supdem_cy <- supdem1 %>%                                   # 1376 obs
    select(country, year, project) %>% 
    distinct()

claassen_replication_input <- map(claassen_input_raw, function(df) {
    claassen_input_cy <- df %>%           # 1868 obs
        mutate(p_dcpo = str_extract(survey, "^[a-z]+"), 
               project = case_when(p_dcpo == "afrob" ~ "afb",
                                   p_dcpo == "amb" ~ "lapop",
                                   p_dcpo == "arabb" ~ "arb",
                                   p_dcpo == "asiab" ~ "asb",
                                   p_dcpo == "asianb" ~ "asnb",
                                   p_dcpo == "neb" ~ "ndb",
                                   p_dcpo == "sasianb" ~ "sab",
                                   TRUE ~ p_dcpo),
               y_dcpo = year) %>%
        select(country, year, y_dcpo, survey, project) %>% 
        unique()

    no_problems <- inner_join(supdem_cy, claassen_input_cy,
                              by = join_by(country, year, project))     # 1283 obs

    needed <- anti_join(supdem_cy, claassen_input_cy,
                        by = join_by(country, year, project))           # 93 obs

    available <- anti_join(claassen_input_cy, supdem_cy,
                           by = join_by(country, year, project))        # 585 obs

    year_fixes <- left_join(needed, available, 
                            by = c("country", "project"),
                            relationship = "many-to-many") %>% # 89 obs
        mutate(diff = year.x - year.y) %>% 
        group_by(country, project, year.x) %>% 
        mutate(closest_to_claassen = min(abs(diff))) %>% 
        ungroup() %>% 
        group_by(country, project, year.y) %>% 
        mutate(closest_to_dcpo = min(abs(diff))) %>% 
        ungroup() %>% 
        filter(closest_to_claassen == abs(diff) & closest_to_dcpo == abs(diff) & abs(diff) <= 3) %>% 
        filter(!(country == "Egypt" & year.x == 2014 & survey == "afrob5")) # double match (it's really afrob6)

    cys_crosswalk <- year_fixes %>% 
        select(country, y_dcpo, y_claassen = year.x, survey)

    missing_cyps <- anti_join(needed, year_fixes,  
                              by = c("country", "year" = "year.x", "project")) # 4 obs; listed in issue #5 (Bahrain now in) plus two unrepresentative Pakistan surveys

    cys_to_drop <- anti_join(available, year_fixes, 
                             by = c("country",
                                    "year" = "year.y",
                                    "project")) %>% # 496 obs
        select(-y_dcpo)

    cri <- df %>% #3705
        remove_bad_items() %>% 
        mutate(p_dcpo = str_extract(survey, "^[a-z]+"), 
               project = case_when(p_dcpo == "afrob" ~ "afb",
                                   p_dcpo == "amb" ~ "lapop",
                                   p_dcpo == "arabb" ~ "arb",
                                   p_dcpo == "asiab" ~ "asb",
                                   p_dcpo == "asianb" ~ "asnb",
                                   p_dcpo == "neb" ~ "ndb",
                                   p_dcpo == "sasianb" ~ "sab",
                                   TRUE ~ p_dcpo),
               i_dcpo = item,
               i_claassen = paste(str_replace(item, "_.*", "_"), p_dcpo)) %>% 
        left_join(cys_crosswalk, by = c("country", "year" = "y_dcpo", "survey")) %>% 
        anti_join(cys_to_drop, by = c("country", "year", "survey")) %>% # surveys unused by Claassen
        mutate(year = if_else(!is.na(y_claassen), y_claassen, year)) %>% # use Claassen's year codings
        mutate(item = if_else(item == "strong_amb_1" & year == 2004, "strong_amb_2", item)) %>% # items conflated in amb_combo file
        with_min_yrs(2) %>% 
        # DCPOtools::format_claassen() %>% 
        # pluck("data") %>% 
        mutate(p_dcpo = str_extract(survey, "^[a-z]+"), 
               project = case_when(p_dcpo == "afrob" ~ "afb",
                                   p_dcpo == "amb" ~ "lapop",
                                   p_dcpo == "arabb" ~ "arb",
                                   p_dcpo == "asiab" ~ "asb",
                                   p_dcpo == "asianb" ~ "asnb",
                                   p_dcpo == "neb" ~ "ndb",
                                   p_dcpo == "sasianb" ~ "sab",
                                   TRUE ~ p_dcpo),
               i_dcpo = item,
               i_claassen0 = paste0(str_replace(item, "_.*", "_"), project),
               i_claassen = case_when(i_claassen0 == "army_afb" ~ "army_afrob",
                                      i_claassen0 == "strong_afb" ~ "strong_afrob",
                                      i_claassen0 == "threestate_afb" ~ "threestate_afrob",
                                      i_claassen0 == "party_afb" ~ "party_afrob",
                                      i_claassen0 == "election_sab" ~ "elec_sab",
                                      TRUE ~ i_claassen0)) %>%
        inner_join(supdem1,
                   by = join_by(country, year, i_claassen)) %>% 
        select(country, year, item = item.x, r, n, survey) %>% 
        format_claassen()

    return(cri)
})

# each element of claassen_replication_input0 is missing 11 country-year-item observations included in supdem1 (6 are apparently no longer available, 2 seem to have never existed, and 3 are unrepresentative and should not have ever been included) and it has 1 more survey item (Claassen conflates two Arabbarometer items with different response categories)

rio::export(claassen_replication_input, 
            file = here("data_replication",
                        "claassen_replication_input.rds"))
```

# Estimating Democratic Support
Below, the inputs created above are used to generate the latent variable estimates of democratic support.  This requires a working installation of [`cmdstanr`](https://mc-stan.org/cmdstanr/index.html).  Generating each set of estimates takes 5 to 10 minutes to run on a M1 MacBook Pro.

## Original sample
```{r claassen_replication_output}
claassen_replication_input <- rio::import(here("data_replication", "claassen_replication_input.rds"))

if (!file.exists(here("data_replication", "supdem.stan.mod5.stan"))) {
    tempfile <- dataverse::get_file_by_doi("doi:10.7910/DVN/HWLW0J/MKBL9E") # AJPS replication file

    writeBin(tempfile, here("data_replication", "supdem.stan.mod5.stan"))
}

cm5 <- cmdstan_model(here("data_replication", "supdem.stan.mod5.stan"))

warmup <- 250; sampling = 75

plan(multisession, workers = 12)
future_options <- furrr_options(seed = 324)

claassen_replication_output <- future_map(claassen_replication_input,
                                          \(cri_df) {
                                              cm5$sample(
                                                  data = cri_df, 
                                                  max_treedepth = 14,
                                                  adapt_delta = 0.99,
                                                  step_size = 0.005,
                                                  seed = 324, 
                                                  chains = 3, 
                                                  parallel_chains = 3,
                                                  iter_warmup = warmup,
                                                  iter_sampling = sampling,
                                                  refresh = warmup/25
                                              )
                                          },
                                          .options = future_options
)

dat <- claassen_replication_input[["mar"]]$data %>% 
    mutate(kk = claassen_replication_input[["mar"]]$jj,
           tt = claassen_replication_input[["mar"]]$tt)

kcodes <- dat %>%
    group_by(country) %>%
    summarize(kk = first(kk) %>% 
                         as.numeric())

tcodes <- dat %>%
    group_by(year) %>%
    summarize(tt = first(tt))

ktcodes <- dat %>%
    group_by(country) %>%
    summarize(first_yr = min(year), 
                     last_yr = max(year))

claassen_replication_theta <- 
    imap(claassen_replication_output,
        \(res, idx) {
            res$draws("theta",
                      format = "df") %>% 
                pivot_longer(starts_with("theta"), values_to = "theta") %>% 
                mutate(tt = as.numeric(gsub("theta\\[(\\d+),\\d+\\]", 
                                            "\\1", name)), 
                       kk = as.numeric(gsub("theta\\[\\d+,(\\d+)\\]", 
                                            "\\1", name))) %>%
                left_join(kcodes, 
                          by = "kk") %>%
                left_join(tcodes, by = "tt") %>%
                mutate(year = if_else(tt == 1,
                                      as.integer(year),
                                      as.integer(min(year, na.rm = TRUE)
                                                 + tt - 1)),
                       set = idx) %>%
                left_join(ktcodes, by = "country") %>%
                arrange(kk, tt) %>%
                group_by(kk, tt) %>% 
                mutate(draw = case_when(idx == "mar" ~ 0,
                                        idx == "lower" ~ n(),
                                        idx == "upper" ~ 2*n(),
                                        idx == "scared" ~ 3*n())
                           + 1:n()) %>% 
                ungroup() %>% 
                select(country, year, theta, set, draw) 
        }) %>% 
    list_rbind() %>% 
    group_split(draw)

saveRDS(claassen_replication_theta, 
        file = here("data_replication", "claassen_replication_theta.rds"))
```

## Expanded 'More Data' Sample
```{r exp_output}
exp_claassen_input <- rio::import(here("data_replication", "exp_claassen_input.rds"))

cm5 <- cmdstan_model(here("data_replication", "supdem.stan.mod5.stan"))

warmup <- 250; sampling = 75

plan(multisession, workers = 12)
future_options <- furrr_options(seed = 324)

exp_output <- future_map(exp_claassen_input,
                         \(cri_df) {
                             cm5$sample(
                                 data = cri_df, 
                                 max_treedepth = 14,
                                 adapt_delta = 0.99,
                                 step_size = 0.005,
                                 seed = 324, 
                                 chains = 3, 
                                 parallel_chains = 3,
                                 iter_warmup = warmup,
                                 iter_sampling = sampling,
                                 refresh = warmup/25
                             )
                         },
                         .options = future_options
)

dat <- exp_claassen_input[["mar"]]$data %>% 
    mutate(kk = exp_claassen_input[["mar"]]$jj,
           tt = exp_claassen_input[["mar"]]$tt)

kcodes <- dat %>%
    group_by(country) %>%
    summarize(kk = first(kk) %>% 
                         as.numeric())

tcodes <- dat %>%
    group_by(year) %>%
    summarize(tt = first(tt))

ktcodes <- dat %>%
    group_by(country) %>%
    summarize(first_yr = min(year), 
                     last_yr = max(year))

exp_theta <- 
    imap(exp_output,
        \(res, idx) {
            res$draws("theta",
                      format = "df") %>% 
                pivot_longer(starts_with("theta"), values_to = "theta") %>% 
                mutate(tt = as.numeric(gsub("theta\\[(\\d+),\\d+\\]", 
                                            "\\1", name)), 
                       kk = as.numeric(gsub("theta\\[\\d+,(\\d+)\\]", 
                                            "\\1", name))) %>%
                left_join(kcodes, 
                          by = "kk") %>%
                left_join(tcodes, by = "tt") %>%
                mutate(year = if_else(tt == 1,
                                      as.integer(year),
                                      as.integer(min(year, na.rm = TRUE)
                                                 + tt - 1)),
                       set = idx) %>%
                left_join(ktcodes, by = "country") %>%
                arrange(kk, tt) %>%
                group_by(kk, tt) %>% 
                mutate(draw = case_when(idx == "mar" ~ 0,
                                        idx == "lower" ~ n(),
                                        idx == "upper" ~ 2*n(),
                                        idx == "scared" ~ 3*n())
                           + 1:n()) %>% 
                ungroup() %>% 
                select(country, year, theta, set, draw) 
        }) %>% 
    list_rbind() %>% 
    group_split(draw)

saveRDS(exp_theta, 
        file = here("data_replication", "exp_theta.rds"))
```

# Merging Data

The output of the next four chunks are the four files used in the (1) 'uncertainty' AJPS replication, (2) 'uncertainty' APSR replication, (3) 'uncertainty & more data AJPS replication', and (4) 'uncertainty & more data' APSR replication.

## AJPS replication, Claassen sample
```{r correct_cls_ajps}
# If latent-variable estimates were not just generated, load them now
if (!exists("claassen_replication_theta")) {
    claassen_replication_theta <- readRDS(file = here("data_replication", "claassen_replication_theta.rds"))
}

# Download controls from osf if file not present
if (!file.exists(here("data_replication",
                      "exp_apsr_cntrl_list.rds"))) { 
    osf_retrieve_file("hf2t4") %>% 
        osf_download(here("data_replication"))
}

cls_ajps_cntrl_list <- readRDS(here("data_replication", "cls_ajps_cntrl_list.rds"))

correct_cls_ajps <- map2(cls_ajps_cntrl_list,
                         claassen_replication_theta,
                         \(controls, theta) {
                             controls %>% 
                                 left_join(theta, by = c("country", "year")) %>% 
                                 mutate(SupDem_trim = ifelse(year < firstyear,
                                                             NA,
                                                             theta),
                                        theta_dem_trim = case_when(
                                            is.na(SupDem_trim) ~ NA,
                                            Regime_VD > 1 & !is.na(SupDem_trim) ~ theta,
                                            TRUE ~ 0),
                                        theta_aut_trim = case_when(
                                            is.na(SupDem_trim) ~ NA,
                                            Regime_VD <= 1 & !is.na(SupDem_trim) ~ theta,
                                            TRUE ~ 0)
                                        ) %>%
                                 select(country, year, theta, contains("trim"),
                                        everything())
                         })

dir.create("data", showWarnings = FALSE)
save(correct_cls_ajps, file = here("data","correct_cls_ajps.rda"))
```

## APSR replication, Claassen sample
```{r correct_cls_apsr}
# If latent-variable estimates were not just generated, load them now
if (!exists("claassen_replication_theta")) {
    claassen_replication_theta <- readRDS(file = here("data_replication", "claassen_replication_theta.rds"))
}

# Download controls from osf if file not present
if (!file.exists(here("data_replication",
                      "exp_apsr_cntrl_list.rds"))) { 
    osf_retrieve_file("fsk96") %>% 
        osf_download(here("data_replication"))
}

cls_apsr_cntrl_list <- readRDS(here("data_replication", "cls_apsr_cntrl_list.rds"))

correct_cls_apsr <- map2(cls_apsr_cntrl_list,
                         claassen_replication_theta,
                         \(controls, theta) {
    controls %>% 
        left_join(theta, by = c("year", "country"))  %>% 
        mutate(SupDem_trim = ifelse(year < firstyear, NA, theta)) %>%
        select(country, year, firstyear, theta, SupDem_trim,
               contains("z"), everything())
})

save(correct_cls_apsr, file = here("data","correct_cls_apsr.rda"))
```

## AJPS replication, expanded sample
```{r expcor_cls_ajps}
# If latent-variable estimates were not just generated, load them now
if (!exists("exp_theta")) {
    exp_theta <- readRDS(file = here("data_replication", "exp_theta.rds"))
}

# Download controls from osf if file not present
if (!file.exists(here("data_replication",
                      "exp_ajps_cntrl_list.rds"))) {
    osf_retrieve_file("etp7a") %>% 
        osf_download(here("data_replication"))
}

exp_ajps_cntrl_list <- readRDS(here("data_replication", "exp_ajps_cntrl_list.rds"))

expcor_cls_ajps <- map2(exp_ajps_cntrl_list,
                 exp_theta,
                 \(controls, theta) {
                     controls %>% 
                         left_join(theta, by = c("country", "year")) %>% 
                         mutate(SupDem_trim = ifelse(year < firstyear, NA, theta),
                                theta_dem_trim = case_when(
                                    is.na(SupDem_trim) ~ NA,
                                    Regime_VD > 1 & !is.na(SupDem_trim) ~ theta,
                                    TRUE ~ 0),
                                theta_aut_trim = case_when(
                                    is.na(SupDem_trim) ~ NA,
                                    Regime_VD <= 1 & !is.na(SupDem_trim) ~ theta,
                                    TRUE ~ 0)) %>%
                         select(country, year, theta, contains("trim"), everything())
                 })

save(expcor_cls_ajps, file = here("data",
                                  "expcor_cls_ajps.rda"))
```

## APSR replication, expanded sample
```{r expcor_cls_apsr}
# If latent-variable estimates were not just generated, load them now
if (!exists("exp_theta")) {
    exp_theta <- readRDS(file = here("data_replication", "exp_theta.rds"))
}

# Download from osf if file not present
if (!file.exists(here("data_replication",
                      "exp_apsr_cntrl_list.rds"))) { 
    osf_retrieve_file("kqtyv") %>% 
        osf_download(here("data_replication"))
}

exp_apsr_cntrl_list <- readRDS(here("data_replication", "exp_apsr_cntrl_list.rds"))

expcor_cls_apsr <- map2(exp_apsr_cntrl_list,
                        claassen_replication_theta,
                        \(controls, theta) {
                            controls %>% 
                                left_join(theta, by = c("year", "country"))  %>% 
                                mutate(SupDem_trim = ifelse(year < firstyear,
                                                            NA, 
                                                            theta)) %>%
                                select(country, year, firstyear, theta, SupDem_trim,
                                       contains("z"), everything())
                        })

save(expcor_cls_apsr, file = here("data",
                                  "expcor_cls_apsr.rda"))
```

# Estimating Democratic Support with the DCPO Model (OSM D)

Online Supplementary Materials D presents, for comparison, the 'Uncertainty & More Data' models when the superior DCPO model is used to estimate the latent variable of democratic support.

## Expanded 'More Data' Sample, DCPO Estimates
```{r dcpo_exp_output}
exp_dcpo_input <- format_dcpo(claassen_input_raw1,
                              scale_q = "threestate_lb",
                              scale_cp = 2)

if (!file.exists(here("data_replication", "dcpo.stan"))) {
    download.file("https://raw.githubusercontent.com/fsolt/DCPO/master/inst/stan/dcpo.stan", here("data_replication", "dcpo.stan"))
}

dcpo <- cmdstan_model(here("data_replication", "dcpo.stan"))

warmup <- 250; sampling = 75

plan(multisession, workers = 12)
future_options <- furrr_options(seed = 324)

exp_dcpo_output <- future_map(exp_dcpo_input,
                         \(edi_df) {
                             dcpo$sample(
                                 data = edi_df[1:14], 
                                 max_treedepth = 14,
                                 adapt_delta = 0.99,
                                 step_size = 0.005,
                                 seed = 324, 
                                 chains = 3, 
                                 parallel_chains = 3,
                                 iter_warmup = warmup,
                                 iter_sampling = sampling,
                                 refresh = warmup/25
                             )
                         },
                         .options = future_options
)

dat <- exp_dcpo_input[["mar"]]$data %>% 
    mutate(kk = exp_dcpo_input[["mar"]]$kk,
           tt = exp_dcpo_input[["mar"]]$tt)

kcodes <- dat %>%
    group_by(country) %>%
    summarize(kk = first(kk) %>% 
                         as.numeric())

tcodes <- dat %>%
    group_by(year) %>%
    summarize(tt = first(tt))

ktcodes <- dat %>%
    group_by(country) %>%
    summarize(first_yr = min(year), 
                     last_yr = max(year))

exp_dcpo_theta <- 
    imap(exp_dcpo_output,
        \(res, idx) {
            res$draws("theta",
                      format = "df") %>% 
                pivot_longer(starts_with("theta"), values_to = "theta") %>% 
                mutate(tt = as.numeric(gsub("theta\\[(\\d+),\\d+\\]", 
                                            "\\1", name)), 
                       kk = as.numeric(gsub("theta\\[\\d+,(\\d+)\\]", 
                                            "\\1", name))) %>%
                left_join(kcodes, 
                          by = "kk") %>%
                left_join(tcodes, by = "tt") %>%
                mutate(year = if_else(tt == 1,
                                      as.integer(year),
                                      as.integer(min(year, na.rm = TRUE)
                                                 + tt - 1)),
                       set = idx) %>%
                left_join(ktcodes, by = "country") %>%
                arrange(kk, tt) %>%
                filter(year >= first_yr & year <= last_yr) %>%
                arrange(kk, tt) %>%
                group_by(kk, tt) %>% 
                mutate(set = idx,
                       draw = case_when(idx == "mar" ~ 0,
                                        idx == "lower" ~ n(),
                                        idx == "upper" ~ 2*n(),
                                        idx == "scared" ~ 3*n())
                           + 1:n()) %>% 
                ungroup() %>% 
                select(country, year, theta, set, draw) 
        }) %>% 
    list_rbind() %>% 
    group_split(draw)

saveRDS(exp_dcpo_theta, 
        file = here("data_replication", "exp_dcpo_theta.rds"))
```

## AJPS replication, DCPO model, expanded sample
```{r expcor_cls_ajps}
# If latent-variable estimates were not just generated, load them now
if (!exists("exp_dcpo_theta")) {
    exp_dcpo_theta <- readRDS(file = here("data_replication", "exp_dcpo_theta.rds"))
}
# If controls not present already, load them now
if (!exists("exp_ajps_cntrl_list")) {
    # download from osf if needed
    if (!file.exists(here("data_replication",
                          "exp_ajps_cntrl_list.rds"))) {
        osf_retrieve_file("etp7a") %>% 
            osf_download(here("data_replication"))
    }

    exp_ajps_cntrl_list <- readRDS(here("data_replication",
                                        "exp_ajps_cntrl_list.rds"))
}

dcpo_ajps <- map2(exp_ajps_cntrl_list,
                 exp_dcpo_theta,
                 \(controls, theta) {
                     controls %>% 
                         left_join(theta, by = c("country", "year")) %>% 
                         mutate(SupDem_trim = ifelse(year < firstyear,
                                                      NA, 
                                                      theta),
                                theta_dem_trim = case_when(
                                    is.na(SupDem_trim) ~ NA,
                                    Regime_VD > 1 & !is.na(SupDem_trim) ~ theta,
                                    TRUE ~ 0),
                                theta_aut_trim = case_when(
                                    is.na(SupDem_trim) ~ NA,
                                    Regime_VD <= 1 & !is.na(SupDem_trim) ~ theta,
                                    TRUE ~ 0)) %>%
                         select(country, year, theta, contains("trim"), everything())
                 })

save(dcpo_ajps, file = here("data",
                            "dcpo_ajps.rda"))
```

## APSR replication, DCPO model, expanded sample
```{r expcor_cls_apsr}
# If latent-variable estimates were not just generated, load them now
if (!exists("exp_dcpo_theta")) {
    exp_dcpo_theta <- readRDS(file = here("data_replication", "exp_dcpo_theta.rds"))
}
# If controls not present already, load them now 
if (!exists("exp_ajps_cntrl_list")) {
    # download from osf if needed
    if (!file.exists(here("data_replication",
                          "exp_apsr_cntrl_list.rds"))) { 
        osf_retrieve_file("kqtyv") %>% 
            osf_download(here("data_replication"))
    }

    exp_apsr_cntrl_list <- readRDS(here("data_replication",
                                        "exp_apsr_cntrl_list.rds"))
}

dcpo_apsr <- map2(exp_apsr_cntrl_list,
                  exp_dcpo_theta,
                  \(controls, theta) {
                      controls %>% 
                          left_join(theta, by = c("year", "country"))  %>% 
                          mutate(SupDem_trim = ifelse(year < firstyear,
                                                      NA, 
                                                      theta)) %>%
                          select(country, year, firstyear, theta, SupDem_trim,
                                 contains("z"), everything())
                  })

save(dcpo_apsr, file = here("data",
                            "dcpo_apsr.rda"))
```